\section{Vision}

The vision of this project is to make a mobile application where the user can compete against other users in real time.
The application is running on a smartphone, but can use multiple 3rd party systems, to integrate tracking, feedback and control of the run.
The user will be able to run against other users live or their ghost as well as their own ghosts on a certain route.
The user can plot their own route on a website, and a score will be calculated taking into account the topography and length of the given route. 
\\ \\
When starting a live run against another user, the application will search for available users, with a route of around the same score as the one provided by the user. If the user selects to run against their own friends or family their own personal running score will be taking into account, and a handicap will be given to the runner with the lowest score, making the competition more even.
\\ \\
Making use of a 3rd party system can enhance the users experience with the application. The 3rd party systems can be pulse watches, headsets for audio feedback, microphone for audio control etc. This can give a more in-depth feeling of running against another player, or tracking your own running data while running.
To ensure that the user is actually running, the application will make use of both GPS, for tracking position, and accelerometer for tracking the kind of movement the user is making.
Data analysis can be applied as even more comprehensive anti-cheat solution.

\subsection{Limitation}

This project is limited both by time, and resources. Due to these restrictions the project for this semester will be limited to a smaller area.
This limits the possibility to implement 3rd party systems, and the main focus of the project will be on making sure the matchmaking part of the project works, and not so much on ensuring the route score is correct, or that one can run against friends and family.
